\chapter*{Acknowledgments}

The cover for the first edition of the book \emph{Compilers: Principles, Techniques, and Tools} (1985) depicted a Knight and a Red Dragon. The Dragon was meant to represent one of the great challenges of engineers in that era: taming the \emph{Compiler}. The field of Computer Science evolved throughout the decades, and the dragon changed its form many times. Today, computing is ubiquitous. Millions of experts in our field fight daily against an increasing army of dragons. And as any Computer Scientist should, I found the Dragon I needed to battle against. This work describes how I, with the help of my family, girlfriend and friends, conquered the dragon of self-belief.

I thank my mother, Angela, for her unrelenting support throughout the countless months where I refused to believe that I would be able to complete this work. I thank my girlfriend, Jessica, for reminding me that I had surpassed many challenges before this work, and that this would be a difficult, but definite step towards achieving my objectives and dreams. I thank my brother, Giovanni, for lending me his pragmatic views on the processes which would help me achieve my goals in this work. I thank my therapist, Joana, for lending me her eyes to realize that this work was not an unbeatable enemy, but a dragon which I strengthened through my own fears. I thank my advisor, Prof. Luciana Nedel, for allowing me to pursue a dangerous but personal objective as my final academic work. And I thank my friends, especially Alexander, Lucas, Monicque and Lorencetti for the company during this long journey, filled with defeat, anger, despair... But ultimately, blissful accomplishment.